---
title: 'R : Modelling Numeric Responses for Prediction'
author: "John Pauline Pineda"
date: "November 29, 2022"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=15, fig.height=10)
```
# **1. Table of Contents**
|
| This document presents a non-exhaustive list of modelling techniques for predicting numeric responses using various helpful packages in <mark style="background-color: #CCECFF">**R**</mark>.    
|
| Regression learning refers to a predictive modelling problem where numerical continuous responses are predicted for a given sample of input data. These models use the training data set and calculate how to approximate a mapping function from input variables to a continuous output variable. Because a regression predictive model predicts a quantity, the performance of the model is typically reported as an error in those predictions. The algorithms applied in this study (mostly contained in the <mark style="background-color: #CCECFF">**caret**</mark> package) attempt to explore the relationships between the response and predictor variables by finding an optimal function for identifying and distinguishing the data into continuous real values and making predictions of that quantity.
|
##  1.1 Sample Data
|
| The <mark style="background-color: #EEEEEE;color: #FF0000">**Solubility**</mark>  dataset from the  <mark style="background-color: #CCECFF">**AppliedPredictiveModeling**</mark> package was used for this illustrated example.
|
| Preliminary dataset assessment:
|
| **[A]** 1267 rows (observations)
|      **[A.1]** Train Set = 951 observations
|      **[A.2]** Test Set = 316 observations
| 
| **[B]** 229 columns (variables)
|      **[B.1]** 1/229 response = <span style="color: #FF0000">Log_Solubility</span> variable (numeric)
|      **[B.2]** 228/229 predictors = All remaining variables (208/228 factor + 20/228 numeric)
|     
| 
```{r section_1.1, warning=FALSE, message=FALSE}
##################################
# Loading R libraries
##################################
library(AppliedPredictiveModeling)
library(caret)
library(rpart)
library(lattice)
library(dplyr)
library(tidyr)
library(moments)
library(skimr)
library(RANN)
library(pls)
library(corrplot)
library(tidyverse)
library(lares)
library(DMwR)
library(gridExtra)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(stats)
library(nnet)
library(elasticnet)
library(earth)
library(party)
library(kernlab)
library(randomForest)
library(Cubist)

##################################
# Loading source and
# formulating the train set
##################################
data(solubility)
Solubility_Train <- as.data.frame(cbind(solTrainY,solTrainX))
Solubility_Test  <- as.data.frame(cbind(solTestY,solTestX))

##################################
# Performing a general exploration of the train set
##################################
dim(Solubility_Train)
str(Solubility_Train)
summary(Solubility_Train)

##################################
# Performing a general exploration of the test set
##################################
dim(Solubility_Test)
str(Solubility_Test)
summary(Solubility_Test)

##################################
# Formulating a data type assessment summary
##################################
PDA <- Solubility_Train
(PDA.Summary <- data.frame(
  Column.Index=c(1:length(names(PDA))),
  Column.Name= names(PDA), 
  Column.Type=sapply(PDA, function(x) class(x)), 
  row.names=NULL)
)
```

##  1.2 Data Quality Assessment
|
| Data quality assessment:
|
| **[A]** No missing observations noted for any variable.
|
| **[B]** Low variance observed for 127 variables with First.Second.Mode.Ratio>5.
|      **[B.1]-[B.33]** <span style="color: #FF0000">FP013</span> to <span style="color: #FF0000">FP045</span> variables (factor)
|      **[B.34]-[B.45]** <span style="color: #FF0000">FP048</span> to <span style="color: #FF0000">FP059</span> variables (factor)
|      **[B.46]** <span style="color: #FF0000">FP114</span> variable (factor)
|      **[B.47]-[B.50]** <span style="color: #FF0000">FP119</span> to <span style="color: #FF0000">FP122</span> variable (factor)
|      **[B.51]-[B.88]** <span style="color: #FF0000">FP124</span> to <span style="color: #FF0000">FP161</span> variables (factor)
|      **[B.89]-[B.118]** <span style="color: #FF0000">FP172</span> to <span style="color: #FF0000">FP201</span> variables (factor)
|      **[B.119]-[B.124]** <span style="color: #FF0000">FP203</span> to <span style="color: #FF0000">FP208</span> variables (factor)
|      **[B.125]** <span style="color: #FF0000">NumSulfer</span> variable (numeric)
|      **[B.126]** <span style="color: #FF0000">NumChlorine</span> variable (numeric)
|      **[B.127]** <span style="color: #FF0000">NumHalogen</span> variable (numeric)
|
| **[C]** Low variance observed for 4 variables with Unique.Count.Ratio<0.01.
|      **[C.1]** <span style="color: #FF0000">NumDblBonds</span> variable (numeric)
|      **[C.2]** <span style="color: #FF0000">NumNitrogen</span> variable (numeric)
|      **[C.3]** <span style="color: #FF0000">NumSulfer</span> variable (numeric)
|      **[C.4]** <span style="color: #FF0000">NumRings</span> variable (numeric)
|
| **[D]** High skewness observed for 3 variables with Skewness>3 or Skewness<(-3).
|      **[D.1]** <span style="color: #FF0000">NumSulfer</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">NumChlorine</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">HydrophilicFactor</span> variable (numeric)
|
```{r section_1.2, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DQA <- Solubility_Train

##################################
# Formulating an overall data quality assessment summary
##################################
(DQA.Summary <- data.frame(
  Column.Index=c(1:length(names(DQA))),
  Column.Name= names(DQA), 
  Column.Type=sapply(DQA, function(x) class(x)), 
  Row.Count=sapply(DQA, function(x) nrow(DQA)),
  NA.Count=sapply(DQA,function(x)sum(is.na(x))),
  Fill.Rate=sapply(DQA,function(x)format(round((sum(!is.na(x))/nrow(DQA)),3),nsmall=3)),
  row.names=NULL)
)

##################################
# Listing all predictors
##################################
DQA.Predictors <- DQA[,!names(DQA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DQA.Predictors.Numeric <- DQA.Predictors[,-(grep("FP", names(DQA.Predictors)))]

if (length(names(DQA.Predictors.Numeric))>0) {
    print(paste0("There are ",
               (length(names(DQA.Predictors.Numeric))),
               " numeric predictor variable(s)."))
} else {
  print("There are no numeric predictor variables.")
}

##################################
# Listing all factor predictors
##################################
DQA.Predictors.Factor <-as.data.frame(lapply(DQA.Predictors[(grep("FP", names(DQA.Predictors)))],factor))

if (length(names(DQA.Predictors.Factor))>0) {
    print(paste0("There are ",
               (length(names(DQA.Predictors.Factor))),
               " factor predictor variable(s)."))
} else {
  print("There are no factor predictor variables.")
}

##################################
# Formulating a data quality assessment summary for factor predictors
##################################
if (length(names(DQA.Predictors.Factor))>0) {
  
  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = x[!(x %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return("x"),
           return(usm[tabsm == max(tabsm)]))
  }
  
  (DQA.Predictors.Factor.Summary <- data.frame(
  Column.Name= names(DQA.Predictors.Factor), 
  Column.Type=sapply(DQA.Predictors.Factor, function(x) class(x)), 
  Unique.Count=sapply(DQA.Predictors.Factor, function(x) length(unique(x))),
  First.Mode.Value=sapply(DQA.Predictors.Factor, function(x) as.character(FirstModes(x)[1])),
  Second.Mode.Value=sapply(DQA.Predictors.Factor, function(x) as.character(SecondModes(x)[1])),
  First.Mode.Count=sapply(DQA.Predictors.Factor, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Predictors.Factor, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  Unique.Count.Ratio=sapply(DQA.Predictors.Factor, function(x) format(round((length(unique(x))/nrow(DQA.Predictors.Factor)),3), nsmall=3)),
  First.Second.Mode.Ratio=sapply(DQA.Predictors.Factor, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  row.names=NULL)
  )
  
} 

##################################
# Formulating a data quality assessment summary for numeric predictors
##################################
if (length(names(DQA.Predictors.Numeric))>0) {
  
  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = na.omit(x)[!(na.omit(x) %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return(0.00001),
           return(usm[tabsm == max(tabsm)]))
  }
  
  (DQA.Predictors.Numeric.Summary <- data.frame(
  Column.Name= names(DQA.Predictors.Numeric), 
  Column.Type=sapply(DQA.Predictors.Numeric, function(x) class(x)), 
  Unique.Count=sapply(DQA.Predictors.Numeric, function(x) length(unique(x))),
  Unique.Count.Ratio=sapply(DQA.Predictors.Numeric, function(x) format(round((length(unique(x))/nrow(DQA.Predictors.Numeric)),3), nsmall=3)),
  First.Mode.Value=sapply(DQA.Predictors.Numeric, function(x) format(round((FirstModes(x)[1]),3),nsmall=3)),
  Second.Mode.Value=sapply(DQA.Predictors.Numeric, function(x) format(round((SecondModes(x)[1]),3),nsmall=3)),
  First.Mode.Count=sapply(DQA.Predictors.Numeric, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Predictors.Numeric, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  First.Second.Mode.Ratio=sapply(DQA.Predictors.Numeric, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  Minimum=sapply(DQA.Predictors.Numeric, function(x) format(round(min(x,na.rm = TRUE),3), nsmall=3)),
  Mean=sapply(DQA.Predictors.Numeric, function(x) format(round(mean(x,na.rm = TRUE),3), nsmall=3)),
  Median=sapply(DQA.Predictors.Numeric, function(x) format(round(median(x,na.rm = TRUE),3), nsmall=3)),
  Maximum=sapply(DQA.Predictors.Numeric, function(x) format(round(max(x,na.rm = TRUE),3), nsmall=3)),
  Skewness=sapply(DQA.Predictors.Numeric, function(x) format(round(skewness(x,na.rm = TRUE),3), nsmall=3)),
  Kurtosis=sapply(DQA.Predictors.Numeric, function(x) format(round(kurtosis(x,na.rm = TRUE),3), nsmall=3)),
  Percentile25th=sapply(DQA.Predictors.Numeric, function(x) format(round(quantile(x,probs=0.25,na.rm = TRUE),3), nsmall=3)),
  Percentile75th=sapply(DQA.Predictors.Numeric, function(x) format(round(quantile(x,probs=0.75,na.rm = TRUE),3), nsmall=3)),
  row.names=NULL)
  )  
  
}

##################################
# Identifying potential data quality issues
##################################

##################################
# Checking for missing observations
##################################
if ((nrow(DQA.Summary[DQA.Summary$NA.Count>0,]))>0){
  print(paste0("Missing observations noted for ",
               (nrow(DQA.Summary[DQA.Summary$NA.Count>0,])),
               " variable(s) with NA.Count>0 and Fill.Rate<1.0."))
  DQA.Summary[DQA.Summary$NA.Count>0,]
} else {
  print("No missing observations noted.")
}

##################################
# Checking for zero or near-zero variance predictors
##################################
if (length(names(DQA.Predictors.Factor))==0) {
  print("No factor predictors noted.")
} else if (nrow(DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,])),
               " factor variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance factor predictors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,])),
               " numeric variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance numeric predictors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,])),
               " numeric variable(s) with Unique.Count.Ratio<0.01."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,]
} else {
  print("No low variance numeric predictors due to low unique count ratio noted.")
}

##################################
# Checking for skewed predictors
##################################
if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),])>0){
  print(paste0("High skewness observed for ",
  (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),])),
  " numeric variable(s) with Skewness>3 or Skewness<(-3)."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                 as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),]
} else {
  print("No skewed numeric predictors noted.")
}

```

##  1.3 Data Preprocessing

###  1.3.1 Outlier
|
| Outlier data assessment:
|
| **[A]** Outliers noted for 20 variables  with the numeric data visualized through a boxplot including observations classified as suspected outliers using the IQR criterion. The IQR criterion means that all observations above the (75th percentile + 1.5 x IQR) or below the (25th percentile - 1.5 x IQR) are suspected outliers, where IQR is the difference between the third quartile (75th percentile) and first quartile (25th percentile). Outlier treatment for numerical stability remains optional depending on potential model requirements for the subsequent steps.
|      **[A.1]** <span style="color: #FF0000">MolWeight	</span> variable (8 outliers detected)
|      **[A.2]** <span style="color: #FF0000">NumAtoms</span> variable (44 outliers detected)
|      **[A.3]** <span style="color: #FF0000">NumNonHAtoms</span> variable (15 outliers detected)
|      **[A.4]** <span style="color: #FF0000">NumBonds</span> variable (51 outliers detected)
|      **[A.5]** <span style="color: #FF0000">NumNonHBonds</span> variable (18 outliers detected)
|      **[A.6]** <span style="color: #FF0000">NumMultBonds</span> variable (6 outliers detected)
|      **[A.7]** <span style="color: #FF0000">NumRotBonds</span> variable (23 outliers detected)
|      **[A.8]** <span style="color: #FF0000">NumDblBonds</span> variable (3 outliers detected)
|      **[A.9]** <span style="color: #FF0000">NumAromaticBonds</span> variable (35 outliers detected)
|      **[A.10]** <span style="color: #FF0000">NumHydrogen</span> variable (32 outliers detected)
|      **[A.11]** <span style="color: #FF0000">NumCarbon</span> variable (35 outliers detected)
|      **[A.12]** <span style="color: #FF0000">NumNitrogen</span> variable (91 outliers detected)
|      **[A.13]** <span style="color: #FF0000">NumOxygen</span> variable (36 outliers detected)
|      **[A.14]** <span style="color: #FF0000">NumSulfer</span> variable (121 outliers detected)
|      **[A.15]** <span style="color: #FF0000">NumChlorine</span> variable (201 outliers detected)
|      **[A.16]** <span style="color: #FF0000">NumHalogen</span> variable (99 outliers detected)
|      **[A.17]** <span style="color: #FF0000">NumRings</span> variable (4 outliers detected)
|      **[A.18]** <span style="color: #FF0000">HydrophilicFactor</span> variable (53 outliers detected)
|      **[A.19]** <span style="color: #FF0000">SurfaceArea1</span> variable (19 outliers detected)
|      **[A.20]** <span style="color: #FF0000">SurfaceArea2</span> variable (12 outliers detected)
|
```{r section_1.3.1, warning=FALSE, message=FALSE, fig.width=15, fig.height=3}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,-(grep("FP", names(DPA.Predictors)))]

##################################
# Identifying outliers for the numeric predictors
##################################
OutlierCountList <- c()

for (i in 1:ncol(DPA.Predictors.Numeric)) {
  Outliers <- boxplot.stats(DPA.Predictors.Numeric[,i])$out
  OutlierCount <- length(Outliers)
  OutlierCountList <- append(OutlierCountList,OutlierCount)
  OutlierIndices <- which(DPA.Predictors.Numeric[,i] %in% c(Outliers))
  boxplot(DPA.Predictors.Numeric[,i], 
          ylab = names(DPA.Predictors.Numeric)[i], 
          main = names(DPA.Predictors.Numeric)[i],
          horizontal=TRUE)
  mtext(paste0(OutlierCount, " Outlier(s) Detected"))
}

OutlierCountSummary <- as.data.frame(cbind(names(DPA.Predictors.Numeric),(OutlierCountList)))
names(OutlierCountSummary) <- c("NumericPredictors","OutlierCount")
OutlierCountSummary$OutlierCount <- as.numeric(as.character(OutlierCountSummary$OutlierCount))
NumericPredictorWithOutlierCount <- nrow(OutlierCountSummary[OutlierCountSummary$OutlierCount>0,])
print(paste0(NumericPredictorWithOutlierCount, " numeric variable(s) were noted with outlier(s)." ))

##################################
# Gathering descriptive statistics
##################################
(DPA_Skimmed <- skim(DPA.Predictors.Numeric))

###################################
# Verifying the data dimensions
###################################
dim(DPA.Predictors.Numeric)

```

###  1.3.2 Zero and Near-Zero Variance
|
| Zero and near-zero variance data assessment:
|
| **[A]** Low variance noted for 127 variables from the previous data quality assessment using a lower threshold.
|
| **[B]** Low variance noted for 3 variables using a preprocessing summary from the <mark style="background-color: #CCECFF">**caret**</mark> package. The <span style="color: #0000FF">nearZeroVar</span> method using both the <span style="color: #0000FF">freqCut</span> and <span style="color: #0000FF">uniqueCut</span> criteria set at 95/5 and 10, respectively, were applied on the dataset.
|      **[B.1]** <span style="color: #FF0000">FP154</span> variable (factor)
|      **[B.2]** <span style="color: #FF0000">FP199</span> variable (factor)
|      **[B.3]** <span style="color: #FF0000">FP200</span> variable (factor)
|
```{r section_1.3.2, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Gathering descriptive statistics
##################################
(DPA_Skimmed <- skim(DPA))

##################################
# Identifying columns with low variance
###################################
DPA_LowVariance <- nearZeroVar(DPA,
                               freqCut = 95/5,
                               uniqueCut = 10,
                               saveMetrics= TRUE)
(DPA_LowVariance[DPA_LowVariance$nzv,])

if ((nrow(DPA_LowVariance[DPA_LowVariance$nzv,]))==0){
  
  print("No low variance predictors noted.")
  
} else {

  print(paste0("Low variance observed for ",
               (nrow(DPA_LowVariance[DPA_LowVariance$nzv,])),
               " numeric variable(s) with First.Second.Mode.Ratio>4 and Unique.Count.Ratio<0.10."))
  
  DPA_LowVarianceForRemoval <- (nrow(DPA_LowVariance[DPA_LowVariance$nzv,]))
  
  print(paste0("Low variance can be resolved by removing ",
               (nrow(DPA_LowVariance[DPA_LowVariance$nzv,])),
               " numeric variable(s)."))
  
  for (j in 1:DPA_LowVarianceForRemoval) {
  DPA_LowVarianceRemovedVariable <- rownames(DPA_LowVariance[DPA_LowVariance$nzv,])[j]
  print(paste0("Variable ",
               j,
               " for removal: ",
               DPA_LowVarianceRemovedVariable))
  }
  
  DPA %>%
  skim() %>%
  dplyr::filter(skim_variable %in% rownames(DPA_LowVariance[DPA_LowVariance$nzv,]))

  ##################################
  # Filtering out columns with low variance
  #################################
  DPA_ExcludedLowVariance <- DPA[,!names(DPA) %in% rownames(DPA_LowVariance[DPA_LowVariance$nzv,])]
  
  ##################################
  # Gathering descriptive statistics
  ##################################
  (DPA_ExcludedLowVariance_Skimmed <- skim(DPA_ExcludedLowVariance))
}

###################################
# Verifying the data dimensions
###################################
dim(DPA_ExcludedLowVariance)

```

###  1.3.3 Collinearity
|
| High collinearity data assessment:
|
| **[A]** High correlation > 95% were noted for 2 variable pairs as confirmed using the preprocessing summaries from the <mark style="background-color: #CCECFF">**caret**</mark> and <mark style="background-color: #CCECFF">**lares**</mark> packages.
|      **[A.1]** <span style="color: #FF0000">NumNonHAtoms</span> and <span style="color: #FF0000">NumNonHBonds</span> variables (numeric)
|      **[A.2]** <span style="color: #FF0000">NumMultBonds</span> and <span style="color: #FF0000">NumAromaticBonds</span> variables (numeric)
|      **[A.3]** <span style="color: #FF0000">NumAtoms</span> and <span style="color: #FF0000">NumBonds</span> variables (numeric)
|
```{r section_1.3.3, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,-(grep("FP", names(DPA.Predictors)))]

##################################
# Visualizing pairwise correlation between predictors
##################################
DPA_CorrelationTest <- cor.mtest(DPA.Predictors.Numeric,
                       method = "pearson",
                       conf.level = .95)

corrplot(cor(DPA.Predictors.Numeric,
             method = "pearson",
             use="pairwise.complete.obs"), 
         method = "circle",
         type = "upper", 
         order = "original", 
         tl.col = "black", 
         tl.cex = 0.75,
         tl.srt = 90, 
         sig.level = 0.05, 
         p.mat = DPA_CorrelationTest$p,
         insig = "blank")



##################################
# Identifying the highly correlated variables
##################################
DPA_Correlation <-  cor(DPA.Predictors.Numeric, 
                        method = "pearson",
                        use="pairwise.complete.obs")
(DPA_HighlyCorrelatedCount <- sum(abs(DPA_Correlation[upper.tri(DPA_Correlation)]) > 0.95))

if (DPA_HighlyCorrelatedCount == 0) {
  print("No highly correlated predictors noted.")
} else {
  print(paste0("High correlation observed for ",
               (DPA_HighlyCorrelatedCount),
               " pairs of numeric variable(s) with Correlation.Coefficient>0.95."))
  
  (DPA_HighlyCorrelatedPairs <- corr_cross(DPA.Predictors.Numeric,
  max_pvalue = 0.05, 
  top = DPA_HighlyCorrelatedCount,
  rm.na = TRUE,
  grid = FALSE
))
  
}


if (DPA_HighlyCorrelatedCount > 0) {
  DPA_HighlyCorrelated <- findCorrelation(DPA_Correlation, cutoff = 0.95)
  
  (DPA_HighlyCorrelatedForRemoval <- length(DPA_HighlyCorrelated))
  
  print(paste0("High correlation can be resolved by removing ",
               (DPA_HighlyCorrelatedForRemoval),
               " numeric variable(s)."))
  
  for (j in 1:DPA_HighlyCorrelatedForRemoval) {
  DPA_HighlyCorrelatedRemovedVariable <- colnames(DPA.Predictors.Numeric)[DPA_HighlyCorrelated[j]]
  print(paste0("Variable ",
               j,
               " for removal: ",
               DPA_HighlyCorrelatedRemovedVariable))
  }
  
  ##################################
  # Filtering out columns with high correlation
  #################################
  DPA_ExcludedHighCorrelation <- DPA[,-DPA_HighlyCorrelated]
  
  ##################################
  # Gathering descriptive statistics
  ##################################
  (DPA_ExcludedHighCorrelation_Skimmed <- skim(DPA_ExcludedHighCorrelation))

}

###################################
# Verifying the data dimensions
###################################
dim(DPA_ExcludedHighCorrelation)

```

###  1.3.4 Linear Dependencies
|
| Linear dependency data assessment:
|
| **[A]** Linear dependencies noted for 2 subsets of variables using the preprocessing summary from the <mark style="background-color: #CCECFF">**caret**</mark> package applying the <span style="color: #0000FF">findLinearCombos</span> method which utilizes the QR decomposition of a matrix to enumerate sets of linear combinations (if they exist).
|
| **[B]** Subset 1
|      **[B.1]** <span style="color: #FF0000">NumNonHBonds</span> variable (numeric)
|      **[B.2]** <span style="color: #FF0000">NumAtoms</span> variable (numeric)
|      **[B.3]** <span style="color: #FF0000">NumNonHAtoms</span> variable (numeric)
|      **[B.3]** <span style="color: #FF0000">NumBonds</span> variable (numeric)
|
| **[C]** Subset 2
|      **[C.1]** <span style="color: #FF0000">NumHydrogen</span> variable (numeric)
|      **[C.2]** <span style="color: #FF0000">NumAtoms</span> variable (numeric)
|      **[C.3]** <span style="color: #FF0000">NumNonHAtoms</span> variable (numeric)
|
```{r section_1.3.4, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,sapply(DPA.Predictors, is.numeric)]

##################################
# Identifying the linearly dependent variables
##################################
DPA_LinearlyDependent <- findLinearCombos(DPA.Predictors.Numeric)

(DPA_LinearlyDependentCount <- length(DPA_LinearlyDependent$linearCombos))

if (DPA_LinearlyDependentCount == 0) {
  print("No linearly dependent predictors noted.")
} else {
  print(paste0("Linear dependency observed for ",
               (DPA_LinearlyDependentCount),
               " subset(s) of numeric variable(s)."))
  
  for (i in 1:DPA_LinearlyDependentCount) {
    DPA_LinearlyDependentSubset <- colnames(DPA.Predictors.Numeric)[DPA_LinearlyDependent$linearCombos[[i]]]
    print(paste0("Linear dependent variable(s) for subset ",
                 i,
                 " include: ",
                 DPA_LinearlyDependentSubset))
  }
  
}

##################################
# Identifying the linearly dependent variables for removal
##################################

if (DPA_LinearlyDependentCount > 0) {
  DPA_LinearlyDependent <- findLinearCombos(DPA.Predictors.Numeric)
  
  DPA_LinearlyDependentForRemoval <- length(DPA_LinearlyDependent$remove)
  
  print(paste0("Linear dependency can be resolved by removing ",
               (DPA_LinearlyDependentForRemoval),
               " numeric variable(s)."))
  
  for (j in 1:DPA_LinearlyDependentForRemoval) {
  DPA_LinearlyDependentRemovedVariable <- colnames(DPA.Predictors.Numeric)[DPA_LinearlyDependent$remove[j]]
  print(paste0("Variable ",
               j,
               " for removal: ",
               DPA_LinearlyDependentRemovedVariable))
  }
  
  ##################################
  # Filtering out columns with linear dependency
  #################################
  DPA_ExcludedLinearlyDependent <- DPA[,-DPA_LinearlyDependent$remove]
  
  ##################################
  # Gathering descriptive statistics
  ##################################
  (DPA_ExcludedLinearlyDependent_Skimmed <- skim(DPA_ExcludedLinearlyDependent))

}

###################################
# Verifying the data dimensions
###################################
dim(DPA_ExcludedLinearlyDependent)

```

###  1.3.5 Shape Transformation
|
| Data transformation assessment:
|
| **[A]** A number of numeric variables in the dataset were observed to be right-skewed which required shape transformation for data distribution stability. Considering that all numeric variables were strictly positive values, the <span style="color: #0000FF">BoxCox</span> method from the <mark style="background-color: #CCECFF">**caret**</mark> package was used to transform their distributional shapes.
|
```{r section_1.3.5, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,-(grep("FP", names(DPA.Predictors)))]

##################################
# Applying a Box-Cox transformation
##################################
DPA_BoxCox <- preProcess(DPA.Predictors.Numeric, method = c("BoxCox"))
DPA_BoxCoxTransformed <- predict(DPA_BoxCox, DPA.Predictors.Numeric)

##################################
# Gathering descriptive statistics
##################################
(DPA_BoxCoxTransformedSkimmed <- skim(DPA_BoxCoxTransformed))

###################################
# Verifying the data dimensions
###################################
dim(DPA_BoxCoxTransformed)

```

###  1.3.6 Centering and Scaling
|
| Centering and scaling data assessment:
|
| **[A]** To maintain numerical stability during modelling, centering and scaling transformations were applied on the transformed numeric variables. The <span style="color: #0000FF">center</span> method from the <mark style="background-color: #CCECFF">**caret**</mark> package was implemented which subtracts the average value of a numeric variable to all the values. As a result of centering, the variables had zero mean values. In addition, the <span style="color: #0000FF">scale</span> method, also from the <mark style="background-color: #CCECFF">**caret**</mark> package, was applied which performs a center transformation with each value of the variable divided by its standard deviation. Scaling the data coerced the values to have a common standard deviation of one.
|
```{r section_1.3.6, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DPA <- Solubility_Train

##################################
# Listing all predictors
##################################
DPA.Predictors <- DPA[,!names(DPA) %in% c("solTrainY")]

##################################
# Listing all numeric predictors
##################################
DPA.Predictors.Numeric <- DPA.Predictors[,-(grep("FP", names(DPA.Predictors)))]

##################################
# Applying a Box-Cox transformation
##################################
DPA_BoxCox <- preProcess(DPA.Predictors.Numeric, method = c("BoxCox"))
DPA_BoxCoxTransformed <- predict(DPA_BoxCox, DPA.Predictors.Numeric)

##################################
# Applying a center and scale data transformation
##################################
DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaled <- preProcess(DPA_BoxCoxTransformed, method = c("center","scale"))
DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed <- predict(DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaled, DPA_BoxCoxTransformed)

##################################
# Gathering descriptive statistics
##################################
(DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformedSkimmed <- skim(DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed))

###################################
# Verifying the data dimensions
###################################
dim(DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed)

```

###  1.3.7 Pre-Processed Dataset
|
| Preliminary dataset assessment:
|
| **[A]** 1267 rows (observations)
|      **[A.1]** Train Set = 951 observations
|      **[A.2]** Test Set = 316 observations
| 
| **[B]** 221 columns (variables)
|      **[B.1]** 1/221 response = <span style="color: #FF0000">Class</span> variable (numeric)
|      **[B.2]** 220/221 predictors = All remaining variables (205/220 factor + 15/220 numeric)
| 
| **[C]** Pre-processing actions applied:
|      **[C.1]** Centering, scaling and shape transformation applied to improve data quality
|      **[C.2]** No outlier treatment applied since the high values noted were contextually valid and sensible 
|      **[C.3]** 3 predictors removed due to zero or near-zero variance 
|      **[C.4]** 3 predictors removed due to high correlation
|      **[C.5]** 2 predictors removed due to linear dependencies
| 
```{r section_1.3.7, warning=FALSE, message=FALSE}
##################################
# Creating the pre-modelling
# train set
##################################
Log_Solubility <- DPA$solTrainY 
PMA.Predictors.Factor   <- DPA.Predictors[,(grep("FP", names(DPA.Predictors)))]
PMA.Predictors.Factor   <- as.data.frame(lapply(PMA.Predictors.Factor,factor))
PMA.Predictors.Numeric  <- DPA.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed
PMA_BoxCoxTransformed_CenteredScaledTransformed <- cbind(Log_Solubility,PMA.Predictors.Factor,PMA.Predictors.Numeric)

##################################
# Filtering out columns noted with data quality issues including
# zero and near-zero variance,
# high correlation and linear dependencies
# to create the pre-modelling dataset
##################################
PMA_BoxCoxTransformed_CenteredScaledTransformed_ExcludedLowVariance_ExcludedLinearlyDependent_ExcludedHighCorrelation <- PMA_BoxCoxTransformed_CenteredScaledTransformed[,!names(PMA_BoxCoxTransformed_CenteredScaledTransformed) %in% c("FP154","FP199","FP200","NumNonHBonds","NumHydrogen","NumNonHAtoms","NumAromaticBonds","NumAtoms")]

PMA_PreModelling_Train <- PMA_BoxCoxTransformed_CenteredScaledTransformed_ExcludedLowVariance_ExcludedLinearlyDependent_ExcludedHighCorrelation

##################################
# Gathering descriptive statistics
##################################
(PMA_PreModelling_Train_Skimmed <- skim(PMA_PreModelling_Train))

###################################
# Verifying the data dimensions
# for the train set
###################################
dim(PMA_PreModelling_Train)

##################################
# Formulating the test set
##################################
DPA_Test <- Solubility_Test
DPA_Test.Predictors <- DPA_Test[,!names(DPA_Test) %in% c("solTestY")]
DPA_Test.Predictors.Numeric <- DPA_Test.Predictors[,-(grep("FP", names(DPA_Test.Predictors)))]
DPA_Test_BoxCox <- preProcess(DPA_Test.Predictors.Numeric, method = c("BoxCox"))
DPA_Test_BoxCoxTransformed <- predict(DPA_Test_BoxCox, DPA_Test.Predictors.Numeric)
DPA_Test.Predictors.Numeric_BoxCoxTransformed_CenteredScaled <- preProcess(DPA_Test_BoxCoxTransformed, method = c("center","scale"))
DPA_Test.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed <- predict(DPA_Test.Predictors.Numeric_BoxCoxTransformed_CenteredScaled, DPA_Test_BoxCoxTransformed)

##################################
# Creating the pre-modelling
# train set
##################################
Log_Solubility <- DPA_Test$solTestY 
PMA_Test.Predictors.Factor   <- DPA_Test.Predictors[,(grep("FP", names(DPA_Test.Predictors)))]
PMA_Test.Predictors.Factor   <- as.data.frame(lapply(PMA_Test.Predictors.Factor,factor))
PMA_Test.Predictors.Numeric  <- DPA_Test.Predictors.Numeric_BoxCoxTransformed_CenteredScaledTransformed
PMA_Test_BoxCoxTransformed_CenteredScaledTransformed <- cbind(Log_Solubility,PMA_Test.Predictors.Factor,PMA_Test.Predictors.Numeric)
PMA_Test_BoxCoxTransformed_CenteredScaledTransformed_ExcludedLowVariance_ExcludedLinearlyDependent_ExcludedHighCorrelation <- PMA_Test_BoxCoxTransformed_CenteredScaledTransformed[,!names(PMA_Test_BoxCoxTransformed_CenteredScaledTransformed) %in% c("FP154","FP199","FP200","NumNonHBonds","NumHydrogen","NumNonHAtoms","NumAromaticBonds","NumAtoms")]

PMA_PreModelling_Test <- PMA_Test_BoxCoxTransformed_CenteredScaledTransformed_ExcludedLowVariance_ExcludedLinearlyDependent_ExcludedHighCorrelation

##################################
# Gathering descriptive statistics
##################################
(PMA_PreModelling_Test_Skimmed <- skim(PMA_PreModelling_Test))

###################################
# Verifying the data dimensions
# for the test set
###################################
dim(PMA_PreModelling_Test)

```

## 1.4 Data Exploration
|
| Exploratory data analysis:
|
| **[A]** Numeric variables which demonstrated linear or non-linear relationships with the <span style="color: #FF0000">Log_Solubility</span> response variable include:
|      **[A.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[A.2]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[A.3]** <span style="color: #FF0000">NumChlorine</span> variable (numeric)
|      **[A.4]** <span style="color: #FF0000">NumHalogen</span> variable (numeric)
|      **[A.5]** <span style="color: #FF0000">NumMultBonds</span> variable (numeric)
|
| **[B]** Factor variables which demonstrated relatively better differentiation of the <span style="color: #FF0000">Log_Solubility</span> response variable between its <span style="color: #FF0000">1</span> and <span style="color: #FF0000">0</span> structure levels include:
|      **[B.1]** <span style="color: #FF0000">FP207</span> variable (factor)
|      **[B.2]** <span style="color: #FF0000">FP190</span> variable (factor)
|      **[B.3]** <span style="color: #FF0000">FP197</span> variable (factor)
|      **[B.4]** <span style="color: #FF0000">FP196</span> variable (factor)
|      **[B.5]** <span style="color: #FF0000">FP193</span> variable (factor)
|      **[B.6]** <span style="color: #FF0000">FP184</span> variable (factor)
|      **[B.7]** <span style="color: #FF0000">FP172</span> variable (factor)
|      **[B.8]** <span style="color: #FF0000">FP149</span> variable (factor)
|      **[B.9]** <span style="color: #FF0000">FP112</span> variable (factor)
|      **[B.10]** <span style="color: #FF0000">FP107</span> variable (factor)
|      **[B.11]** <span style="color: #FF0000">FP089</span> variable (factor)
|      **[B.12]** <span style="color: #FF0000">FP076</span> variable (factor)
|      **[B.13]** <span style="color: #FF0000">FP059</span> variable (factor)
|      **[B.14]** <span style="color: #FF0000">FP049</span> variable (factor)
|      **[B.15]** <span style="color: #FF0000">FP044</span> variable (factor)
|      **[B.16]** <span style="color: #FF0000">FP014</span> variable (factor)
|      **[B.17]** <span style="color: #FF0000">FP013</span> variable (factor)
|
```{r section_1.4, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
EDA <- PMA_PreModelling_Train

##################################
# Listing all predictors
##################################
EDA.Predictors <- EDA[,!names(EDA) %in% c("Log_Solubility")]

##################################
# Listing all numeric predictors
##################################
EDA.Predictors.Numeric <- EDA.Predictors[,sapply(EDA.Predictors, is.numeric)]
ncol(EDA.Predictors.Numeric)
names(EDA.Predictors.Numeric)

##################################
# Listing all factor predictors
##################################
EDA.Predictors.Factor <- EDA.Predictors[,sapply(EDA.Predictors, is.factor)]
ncol(EDA.Predictors.Factor)
names(EDA.Predictors.Factor)

##################################
# Formulating the scatter plots
##################################
featurePlot(x = EDA.Predictors.Numeric, 
            y = EDA$Log_Solubility,
            between = list(x = 1, y = 1),
            type = c("g", "p", "smooth"),
            labels = rep("", 2))

##################################
# Restructuring the dataset for
# for boxplot analysis
##################################
Log_Solubility <- DPA$solTrainY
EDA.Boxplot.Source <- cbind(Log_Solubility,
                     EDA.Predictors.Factor)


EDA.Boxplot.Gathered.Group1 <- gather(EDA.Boxplot.Source,
                                      'FP001','FP002','FP003','FP004','FP005',
                                      'FP006','FP007','FP008','FP009','FP010',
                                      'FP011','FP012','FP013','FP014','FP015',
                                      'FP016','FP017','FP018','FP019','FP020',
                                      'FP021','FP022','FP023','FP024','FP025',
                                      'FP026','FP027','FP028','FP029','FP030',
                                      'FP031','FP032','FP033','FP034','FP035',
                                      'FP036','FP037','FP038','FP039','FP040',
                                      key="Descriptor",
                                      value="Structure")

EDA.Boxplot.Gathered.Group2 <- gather(EDA.Boxplot.Source,
                                      'FP041','FP042','FP043','FP044','FP045',
                                      'FP046','FP047','FP048','FP049','FP050',
                                      'FP051','FP052','FP053','FP054','FP055',
                                      'FP056','FP057','FP058','FP059','FP060',
                                      'FP061','FP062','FP063','FP064','FP065',
                                      'FP066','FP067','FP068','FP069','FP070',
                                      'FP071','FP072','FP073','FP074','FP075',
                                      'FP076','FP077','FP078','FP079','FP080',
                                      key="Descriptor",
                                      value="Structure")

EDA.Boxplot.Gathered.Group3 <- gather(EDA.Boxplot.Source,
                                      'FP081','FP082','FP083','FP084','FP085',
                                      'FP086','FP087','FP088','FP089','FP090',
                                      'FP091','FP092','FP093','FP094','FP095',
                                      'FP096','FP097','FP098','FP099','FP100',
                                      'FP101','FP102','FP103','FP104','FP105',
                                      'FP106','FP107','FP108','FP109','FP110',
                                      'FP111','FP112','FP113','FP114','FP115',
                                      'FP116','FP117','FP118','FP119','FP120',
                                      key="Descriptor",
                                      value="Structure")

EDA.Boxplot.Gathered.Group4 <- gather(EDA.Boxplot.Source,
                                      'FP121','FP122','FP123','FP124','FP125',
                                      'FP126','FP127','FP128','FP129','FP130',
                                      'FP131','FP132','FP133','FP134','FP135',
                                      'FP136','FP137','FP138','FP139','FP140',
                                      'FP141','FP142','FP143','FP144','FP145',
                                      'FP146','FP147','FP148','FP149','FP150',
                                      'FP151','FP152','FP153','FP155','FP156',
                                      'FP157','FP158','FP159','FP160','FP161',
                                      key="Descriptor",
                                      value="Structure")

EDA.Boxplot.Gathered.Group5 <- gather(EDA.Boxplot.Source,
                                      'FP162','FP163','FP164','FP165','FP166',
                                      'FP167','FP168','FP169','FP170','FP171',
                                      'FP172','FP173','FP174','FP175','FP176',
                                      'FP177','FP178','FP179','FP180','FP181',
                                      'FP182','FP183','FP184','FP185','FP186',
                                      'FP187','FP188','FP189','FP190','FP191',
                                      'FP192','FP193','FP194','FP195','FP196',
                                      'FP197','FP198','FP201','FP202','FP203',
                                      'FP204','FP205','FP206','FP207','FP208',
                                      key="Descriptor",
                                      value="Structure")

bwplot(Log_Solubility~Structure|Descriptor,
       data=EDA.Boxplot.Gathered.Group5,
       ylab="Log Solubility",
       xlab="Structure",
       layout=(c(9,5)))

bwplot(Log_Solubility~Structure|Descriptor,
       data=EDA.Boxplot.Gathered.Group4,
       ylab="Log Solubility",
       xlab="Structure",
       layout=(c(9,5)))

bwplot(Log_Solubility~Structure|Descriptor,
       data=EDA.Boxplot.Gathered.Group3,
       ylab="Log Solubility",
       xlab="Structure",
       layout=(c(9,5)))

bwplot(Log_Solubility~Structure|Descriptor,
       data=EDA.Boxplot.Gathered.Group2,
       ylab="Log Solubility",
       xlab="Structure",
       layout=(c(9,5)))

bwplot(Log_Solubility~Structure|Descriptor,
       data=EDA.Boxplot.Gathered.Group1,
       ylab="Log Solubility",
       xlab="Structure",
       layout=(c(9,5)))

```

## 1.5 Predictive Model Development
###  1.5.1 Linear Regression (LR)
|
| **[A]** The linear regression model from the  <mark style="background-color: #CCECFF">**stats**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">intercept</span> = intercept held constant at a value of TRUE
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves intercept=TRUE
|      **[C.2]** Root-Mean-Square Error = 0.68719
|      **[C.3]** R-Squared = 0.88629
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">NumRotBonds</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">FP072 (Structure=1)</span> variable (factor)
|      **[D.4]** <span style="color: #FF0000">NumOxygen</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">NumMultBonds</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.77258
|      **[E.2]** R-Squared = 0.86439
|
```{r section_1.5.1, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_LR <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_LR$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
# No hyperparameter tuning process conducted
# hyperparameter=intercept fixed to TRUE

##################################
# Running the linear regression model
# by setting the caret method to 'lm'
##################################
set.seed(12345678)
LR_Tune <- train(x = PMA_PreModelling_Train_LR[,!names(PMA_PreModelling_Train_LR) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_LR$Log_Solubility,
                 method = "lm",
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
LR_Tune

LR_Tune$finalModel

LR_Tune$results

(LR_Train_RMSE <- LR_Tune$results$RMSE)
(LR_Train_Rsquared <- LR_Tune$results$Rsquared)

##################################
# Identifying and plotting the
# best model predictors
##################################
LR_VarImp <- varImp(LR_Tune, scale = TRUE)
plot(LR_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Linear Regression",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
LR_Test <- data.frame(LR_Observed = PMA_PreModelling_Test$Log_Solubility,
                      LR_Predicted = predict(LR_Tune, 
                                            PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

LR_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(LR_Test_Metrics <- postResample(LR_Test[,2], LR_Test[,1]))
(LR_Test_RMSE <- LR_Test_Metrics[1])
(LR_Test_Rsquared <- LR_Test_Metrics[2])

```

###  1.5.2 Penalized Linear Regression - Ridge (PLR_R)
|
| **[A]** The penalized linear regression (ridge) model from the  <mark style="background-color: #CCECFF">**elasticnet**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">lambda</span> = weight decay made to vary across a range of values equal to 0.00 to 0.10
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves lambda=0.025
|      **[C.2]** Root-Mean-Square Error = 0.65275
|      **[C.3]** R-Squared = 0.89684
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.74148
|      **[E.2]** R-Squared = 0.87517
|
```{r section_1.5.2, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_PLR_R <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_PLR_R)

PMA_PreModelling_Test_PLR_R <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_PLR_R)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_PLR_R$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
PLR_R_Grid = expand.grid(lambda = seq(0, 0.10, length = 5))

##################################
# Running the penalized linear regression (ridge) model
# by setting the caret method to 'ridge'
##################################
set.seed(12345678)
PLR_R_Tune <- train(x = PMA_PreModelling_Train_PLR_R[,!names(PMA_PreModelling_Train_PLR_R) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_PLR_R$Log_Solubility,
                 method = "ridge",
                 tuneGrid = PLR_R_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
PLR_R_Tune

PLR_R_Tune$finalModel

PLR_R_Tune$results

(PLR_R_Train_RMSE <- PLR_R_Tune$results[PLR_R_Tune$results$lambda==PLR_R_Tune$bestTune$lambda,
                 c("RMSE")])
(PLR_R_Train_Rsquared <- PLR_R_Tune$results[PLR_R_Tune$results$lambda==PLR_R_Tune$bestTune$lambda,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
PLR_R_Test <- data.frame(PLR_R_Observed = PMA_PreModelling_Test$Log_Solubility,
                         PLR_R_Predicted = predict(PLR_R_Tune,
                         PMA_PreModelling_Test_PLR_R[,!names(PMA_PreModelling_Test_PLR_R) %in% c("Log_Solubility")]))

PLR_R_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(PLR_R_Test_Metrics <- postResample(PLR_R_Test[,2], PLR_R_Test[,1]))
(PLR_R_Test_RMSE <- PLR_R_Test_Metrics[1])
(PLR_R_Test_Rsquared <- PLR_R_Test_Metrics[2])

```

###  1.5.3 Penalized Linear Regression - Lasso (PLR_L)
|
| **[A]** The penalized linear regression (lasso) model from the  <mark style="background-color: #CCECFF">**elasticnet**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package.
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">fraction</span> = fraction of full solution made to vary across a range of values equal to 0.05 to 1.00
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves fraction=0.525
|      **[C.2]** Root-Mean-Square Error = 0.64896
|      **[C.3]** R-Squared = 0.89763
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.73891
|      **[E.2]** R-Squared = 0.87468
|
```{r section_1.5.3, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_PLR_L <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_PLR_L)

PMA_PreModelling_Test_PLR_L <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_PLR_L)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_PLR_L$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
PLR_L_Grid = expand.grid(fraction = seq(0.05, 1.00, length = 5))

##################################
# Running the penalized linear regression (lasso) model
# by setting the caret method to 'lasso'
##################################
set.seed(12345678)
PLR_L_Tune <- train(x = PMA_PreModelling_Train_PLR_L[,!names(PMA_PreModelling_Train_PLR_L) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_PLR_L$Log_Solubility,
                 method = "lasso",
                 tuneGrid = PLR_L_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
PLR_L_Tune

PLR_L_Tune$finalModel

PLR_L_Tune$results

(PLR_L_Train_RMSE <- PLR_L_Tune$results[PLR_L_Tune$results$fraction==PLR_L_Tune$bestTune$fraction,
                 c("RMSE")])
(PLR_L_Train_Rsquared <- PLR_L_Tune$results[PLR_L_Tune$results$fraction==PLR_L_Tune$bestTune$fraction,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
PLR_L_Test <- data.frame(PLR_L_Observed = PMA_PreModelling_Test$Log_Solubility,
                         PLR_L_Predicted = predict(PLR_L_Tune,
                         PMA_PreModelling_Test_PLR_L[,!names(PMA_PreModelling_Test_PLR_L) %in% c("Log_Solubility")]))

PLR_L_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(PLR_L_Test_Metrics <- postResample(PLR_L_Test[,2], PLR_L_Test[,1]))
(PLR_L_Test_RMSE <- PLR_L_Test_Metrics[1])
(PLR_L_Test_Rsquared <- PLR_L_Test_Metrics[2])

```

###  1.5.4 Penalized Linear Regression - ElasticNet (PLR_E)
|
| **[A]** The penalized linear regression (elasticnet) model from the  <mark style="background-color: #CCECFF">**elasticnet**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 2 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">lambda</span> = weight decay made to vary across a range of values equal to 0.00 to 0.10
|      **[B.2]** <span style="color: #FF0000">fraction</span> = fraction of full solution made to vary across a range of values equal to 0.05 to 1.00
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves lambda=0.010 and fraction=0.7625
|      **[C.2]** Root-Mean-Square Error = 0.64716
|      **[C.3]** R-Squared = 0.89822
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.73519
|      **[E.2]** R-Squared = 0.87618
|
```{r section_1.5.4, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_PLR_E <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_PLR_E)

PMA_PreModelling_Test_PLR_E <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_PLR_E)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_PLR_E$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
PLR_E_Grid = expand.grid(lambda = c(0, 0.01, 0.10), 
                        fraction = seq(0.05, 1.00, length = 5))

##################################
# Running the penalized linear regression (elasticnet) model
# by setting the caret method to 'enet'
##################################
set.seed(12345678)
PLR_E_Tune <- train(x = PMA_PreModelling_Train_PLR_E[,!names(PMA_PreModelling_Train_PLR_E) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_PLR_E$Log_Solubility,
                 method = "enet",
                 tuneGrid = PLR_E_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
PLR_E_Tune

PLR_E_Tune$finalModel

PLR_E_Tune$results

(PLR_E_Train_RMSE <- PLR_E_Tune$results[PLR_E_Tune$results$fraction==PLR_E_Tune$bestTune$fraction &
                              PLR_E_Tune$results$lambda==PLR_E_Tune$bestTune$lambda,
                 c("RMSE")])
(PLR_E_Train_Rsquared <- PLR_E_Tune$results[PLR_E_Tune$results$fraction==PLR_E_Tune$bestTune$fraction &
                              PLR_E_Tune$results$lambda==PLR_E_Tune$bestTune$lambda,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
PLR_E_Test <- data.frame(PLR_E_Observed = PMA_PreModelling_Test$Log_Solubility,
                         PLR_E_Predicted = predict(PLR_E_Tune,
                         PMA_PreModelling_Test_PLR_E[,!names(PMA_PreModelling_Test_PLR_E) %in% c("Log_Solubility")]))

PLR_E_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(PLR_E_Test_Metrics <- postResample(PLR_E_Test[,2], PLR_E_Test[,1]))
(PLR_E_Test_RMSE <- PLR_E_Test_Metrics[1])
(PLR_E_Test_Rsquared <- PLR_E_Test_Metrics[2])

```

###  1.5.5 Principal Component Regression (PCR)
|
| **[A]** The principal component regression model from the  <mark style="background-color: #CCECFF">**pls**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">ncomp</span> = number of components made to vary across a range of values equal to 1 to 35
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves ncomp=34
|      **[C.2]** Root-Mean-Square Error = 0.74260
|      **[C.3]** R-Squared = 0.86772
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.84483
|      **[E.2]** R-Squared = 0.83516
|
```{r section_1.5.5, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_PCR <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_PCR$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
PCR_Grid = expand.grid(ncomp = 1:35)

##################################
# Running the principal component regression model
# by setting the caret method to 'pcr'
##################################
set.seed(12345678)
PCR_Tune <- train(x = PMA_PreModelling_Train_PCR[,!names(PMA_PreModelling_Train_PCR) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_PCR$Log_Solubility,
                 method = "pcr",
                 tuneGrid = PCR_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
PCR_Tune

PCR_Tune$finalModel

PCR_Tune$results

(PCR_Train_RMSE <- PCR_Tune$results[PCR_Tune$results$ncomp==PCR_Tune$bestTune$ncomp,
                 c("RMSE")])
(PCR_Train_Rsquared <- PCR_Tune$results[PCR_Tune$results$ncomp==PCR_Tune$bestTune$ncomp,
                 c("Rsquared")])



##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
PCR_Test <- data.frame(PCR_Observed = PMA_PreModelling_Test$Log_Solubility,
                      PCR_Predicted = predict(PCR_Tune, 
                      PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

PCR_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(PCR_Test_Metrics <- postResample(PCR_Test[,2], PCR_Test[,1]))
(PCR_Test_RMSE <- PCR_Test_Metrics[1])
(PCR_Test_Rsquared <- PCR_Test_Metrics[2])

```

###  1.5.6 Partial Least Squares (PLS)
|
| **[A]** The partial least squares model from the  <mark style="background-color: #CCECFF">**pls**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">ncomp</span> = number of components made to vary across a range of values equal to 1 to 35
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves ncomp=16
|      **[C.1]** Root-Mean-Square Error = 0.64404
|      **[C.2]** R-Squared = 0.89921
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">NumMultBonds</span> variable (numeric)
|      **[D.4]** <span style="color: #FF0000">NumChlorine</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">NumHalogen</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.76473
|      **[E.2]** R-Squared = 0.86706
|
```{r section_1.5.6, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_PLS <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_PLS$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
PLS_Grid = expand.grid(ncomp = 1:35)

##################################
# Running the partial least squares model
# by setting the caret method to 'pls'
##################################
set.seed(12345678)
PLS_Tune <- train(x = PMA_PreModelling_Train_PLS[,!names(PMA_PreModelling_Train_PLS) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_PLS$Log_Solubility,
                 method = "pls",
                 tuneGrid = PLS_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
PLS_Tune

PLS_Tune$finalModel

PLS_Tune$results

(PLS_Train_RMSE <- PLS_Tune$results[PLS_Tune$results$ncomp==PLS_Tune$bestTune$ncomp,
                 c("RMSE")])
(PLS_Train_Rsquared <- PLS_Tune$results[PLS_Tune$results$ncomp==PLS_Tune$bestTune$ncomp,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
PLS_VarImp <- varImp(PLS_Tune, scale = TRUE)
plot(PLS_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Partial Least Squares",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
PLS_Test <- data.frame(PLS_Observed = PMA_PreModelling_Test$Log_Solubility,
                       PLS_Predicted = predict(PLS_Tune, 
                       PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

PLS_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(PLS_Test_Metrics <- postResample(PLS_Test[,2], PLS_Test[,1]))
(PLS_Test_RMSE <- PLS_Test_Metrics[1])
(PLS_Test_Rsquared <- PLS_Test_Metrics[2])

```

###  1.5.7 Averaged Neural Network (AVNN)
|
| **[A]** The averaged neural network model from the  <mark style="background-color: #CCECFF">**nnet**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 3 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">size</span> = number of hidden units made to vary across a range of values equal to 1 to 13
|      **[B.2]** <span style="color: #FF0000">decay</span> = weight decay made to vary across a range of values equal to 0.00 to 0.10
|      **[B.3]** <span style="color: #FF0000">bag</span> = bagging held constant at a value of FALSE
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves size=13, decay=0.01 and bag=FALSE
|      **[C.2]** Root-Mean-Square Error = 1.06088
|      **[C.3]** R-Squared = 0.74802
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.98624
|      **[E.2]** R-Squared = 0.78293
|
```{r section_1.5.7, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_AVNN <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_AVNN)

PMA_PreModelling_Test_AVNN <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_AVNN)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_AVNN$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
AVNN_Grid = expand.grid(decay = c(0.00, 0.01, 0.10), 
                        size = c(1, 5, 9, 13), 
                        bag = FALSE)
maxSize <- max(AVNN_Grid$size)

##################################
# Running the averaged neural network model
# by setting the caret method to 'avNNet'
##################################
set.seed(12345678)
AVNN_Tune <- train(x = PMA_PreModelling_Train_AVNN[,!names(PMA_PreModelling_Train_AVNN) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_AVNN$Log_Solubility,
                 method = "avNNet",
                 tuneGrid = AVNN_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"),
                 linout = TRUE,
                 trace = FALSE,
                 MaxNWts = maxSize * ((ncol(PMA_PreModelling_Train_AVNN)-1) + 1) + maxSize + 1,
                 maxit = 5,
                 allowParallel = FALSE)

##################################
# Reporting the cross-validation results
# for the train set
##################################
AVNN_Tune

AVNN_Tune$finalModel

AVNN_Tune$results

(AVNN_Train_RMSE <- AVNN_Tune$results[AVNN_Tune$results$decay==AVNN_Tune$bestTune$decay &
                              AVNN_Tune$results$size==AVNN_Tune$bestTune$size,
                 c("RMSE")])
(AVNN_Train_Rsquared <- AVNN_Tune$results[AVNN_Tune$results$decay==AVNN_Tune$bestTune$decay &
                              AVNN_Tune$results$size==AVNN_Tune$bestTune$size,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
AVNN_Test <- data.frame(AVNN_Observed = PMA_PreModelling_Test$Log_Solubility,
                         AVNN_Predicted = predict(AVNN_Tune,
                         PMA_PreModelling_Test_AVNN[,!names(PMA_PreModelling_Test_AVNN) %in% c("Log_Solubility")]))

AVNN_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(AVNN_Test_Metrics <- postResample(AVNN_Test[,2], AVNN_Test[,1]))
(AVNN_Test_RMSE <- AVNN_Test_Metrics[1])
(AVNN_Test_Rsquared <- AVNN_Test_Metrics[2])

```

###  1.5.8 Multivariate Adaptive Regression Splines (MARS)
|
| **[A]** The multivariate adaptive regression splines model from the  <mark style="background-color: #CCECFF">**earth**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 2 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">nprune</span> = number of terms made to vary across a range of values equal to 1 to 5
|      **[B.2]** <span style="color: #FF0000">degree</span> = product degree made to vary across a range of values equal to 5 to 20
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves nprune=20 and degree=3
|      **[C.2]** Root-Mean-Square Error = 0.70348
|      **[C.3]** R-Squared = 0.88230
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">SurfaceArea2</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.4]** <span style="color: #FF0000">SurfaceArea1</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">FP142 (Structure=1)</span> variable (factor)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.75804
|      **[E.2]** R-Squared = 0.86891
|
```{r section_1.5.8, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_MARS <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_MARS$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
MARS_Grid = expand.grid(expand.grid(degree = 1:5, nprune = seq(5, 20, length = 4)))

##################################
# Running the multivariate adaptive regression splines model
# by setting the caret method to 'earth'
##################################
set.seed(12345678)
MARS_Tune <- train(x = PMA_PreModelling_Train_MARS[,!names(PMA_PreModelling_Train_MARS) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_MARS$Log_Solubility,
                 method = "earth",
                 tuneGrid = MARS_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
MARS_Tune

MARS_Tune$finalModel

MARS_Tune$results

(MARS_Train_RMSE <- MARS_Tune$results[MARS_Tune$results$nprune==MARS_Tune$bestTune$nprune &
                              MARS_Tune$results$degree==MARS_Tune$bestTune$degree,
                 c("RMSE")])
(MARS_Train_Rsquared <- MARS_Tune$results[MARS_Tune$results$nprune==MARS_Tune$bestTune$nprune &
                              MARS_Tune$results$degree==MARS_Tune$bestTune$degree,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
MARS_VarImp <- varImp(MARS_Tune, scale = TRUE)
plot(MARS_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Multivariate Adaptive Regression Splines",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
MARS_Test <- data.frame(MARS_Observed = PMA_PreModelling_Test$Log_Solubility,
                      MARS_Predicted = predict(MARS_Tune, 
                                            PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

MARS_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(MARS_Test_Metrics <- postResample(MARS_Test[,2], MARS_Test[,1]))
(MARS_Test_RMSE <- MARS_Test_Metrics[1])
(MARS_Test_Rsquared <- MARS_Test_Metrics[2])

```

###  1.5.9 Support Vector Machine - Radial Basis Function Kernel (SVM_R)
|
| **[A]** The support vector machine (radial basis function kernel) model from the  <mark style="background-color: #CCECFF">**kernlab**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 2 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">sigma</span> = sigma held constant at a value of 0.00285
|      **[B.2]** <span style="color: #FF0000">C</span> = cost made to vary across a range of 14 default values
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves sigma=0.00285 and C=16
|      **[C.2]** Root-Mean-Square Error = 0.59505
|      **[C.3]** R-Squared = 0.91551
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.62742
|      **[E.2]** R-Squared = 0.90989
|
```{r section_1.5.9, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_SVM_R <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_SVM_R)

PMA_PreModelling_Test_SVM_R <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_SVM_R)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_SVM_R$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
# used a range of default values

##################################
# Running the support vector machine (radial basis function kernel) model
# by setting the caret method to 'svmRadial'
##################################
set.seed(12345678)
SVM_R_Tune <- train(x = PMA_PreModelling_Train_SVM_R[,!names(PMA_PreModelling_Train_SVM_R) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_SVM_R$Log_Solubility,
                 method = "svmRadial",
                 tuneLength = 14,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
SVM_R_Tune

SVM_R_Tune$finalModel

SVM_R_Tune$results

SVM_R_Train_RMSE <- SVM_R_Tune$results[SVM_R_Tune$results$C==SVM_R_Tune$bestTune$C,
                 c("RMSE")]
SVM_R_Train_Rsquared <- SVM_R_Tune$results[SVM_R_Tune$results$C==SVM_R_Tune$bestTune$C,
                 c("Rsquared")]

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
SVM_R_Test <- data.frame(SVM_R_Observed = PMA_PreModelling_Test$Log_Solubility,
                         SVM_R_Predicted = predict(SVM_R_Tune,
                         PMA_PreModelling_Test_SVM_R[,!names(PMA_PreModelling_Test_SVM_R) %in% c("Log_Solubility")]))

SVM_R_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(SVM_R_Test_Metrics <- postResample(SVM_R_Test[,2], SVM_R_Test[,1]))
(SVM_R_Test_RMSE <- SVM_R_Test_Metrics[1])
(SVM_R_Test_Rsquared <- SVM_R_Test_Metrics[2])

```

###  1.5.10 Support Vector Machine - Polynomial Kernel (SVM_P)
|
| **[A]** The support vector machine (polynomial kernel) model from the  <mark style="background-color: #CCECFF">**kernlab**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 3 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">degree</span> = polynomial degree made to vary across a range of values equal to 1 to 2
|      **[B.2]** <span style="color: #FF0000">scale</span> = scale made to vary across a range of values equal to 0.001 to 0.010
|      **[B.3]** <span style="color: #FF0000">C</span> = cost made to vary across a range of values equal to 0.25 to 32.00
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves degree=2, scale=0.001 and C=8
|      **[C.2]** Root-Mean-Square Error = 0.60281
|      **[C.3]** R-Squared = 0.91167
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.63778
|      **[E.2]** R-Squared = 0.90623
|
```{r section_1.5.10, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_SVM_P <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_SVM_P)

PMA_PreModelling_Test_SVM_P <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_SVM_P)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_SVM_P$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
SVM_P_Grid = expand.grid(degree = 1:2, 
                       scale = c(0.01, 0.005, 0.001), 
                       C = 2^(-2:5))

##################################
# Running the support vector machine (polynomial kernel) model
# by setting the caret method to 'svmPoly'
##################################
set.seed(12345678)
SVM_P_Tune <- train(x = PMA_PreModelling_Train_SVM_P[,!names(PMA_PreModelling_Train_SVM_P) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_SVM_P$Log_Solubility,
                 method = "svmPoly",
                 tuneGrid = SVM_P_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
SVM_P_Tune

SVM_P_Tune$finalModel

SVM_P_Tune$results

(SVM_P_Train_RMSE <- SVM_P_Tune$results[SVM_P_Tune$results$degree==SVM_P_Tune$bestTune$degree &
                              SVM_P_Tune$results$scale==SVM_P_Tune$bestTune$scale &
                              SVM_P_Tune$results$C==SVM_P_Tune$bestTune$C,
                 c("RMSE")])
(SVM_P_Train_Rsquared <- SVM_P_Tune$results[SVM_P_Tune$results$degree==SVM_P_Tune$bestTune$degree &
                              SVM_P_Tune$results$scale==SVM_P_Tune$bestTune$scale &
                              SVM_P_Tune$results$C==SVM_P_Tune$bestTune$C,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
SVM_P_Test <- data.frame(SVM_P_Observed = PMA_PreModelling_Test$Log_Solubility,
                         SVM_P_Predicted = predict(SVM_P_Tune,
                         PMA_PreModelling_Test_SVM_P[,!names(PMA_PreModelling_Test_SVM_P) %in% c("Log_Solubility")]))

SVM_P_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(SVM_P_Test_Metrics <- postResample(SVM_P_Test[,2], SVM_P_Test[,1]))
(SVM_P_Test_RMSE <- SVM_P_Test_Metrics[1])
(SVM_P_Test_Rsquared <- SVM_P_Test_Metrics[2])

```

###  1.5.11 K-Nearest Neighbors (KNN)
|
| **[A]** The k-nearest neighbors model was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">k</span> = number of neighbors made to vary across a range of values equal to 1 to 15
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves k=3
|      **[C.2]** Root-Mean-Square Error = 1.06729
|      **[C.3]** R-Squared = 0.73260
|
| **[D]** The model does not allow for ranking of predictors in terms of variable importance.
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 1.12471
|      **[E.2]** R-Squared = 0.71373
|
```{r section_1.5.11, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_KNN <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_KNN)

PMA_PreModelling_Test_KNN <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_KNN)

##################################
# Creating consistent fold assignments
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_KNN$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
KNN_Grid = data.frame(k = 1:15)

##################################
# Running the k-nearest neighbors model
# by setting the caret method to 'knn'
##################################
set.seed(12345678)
KNN_Tune <- train(x = PMA_PreModelling_Train_KNN[,!names(PMA_PreModelling_Train_KNN) %in% c("Log_Solubility")],
                 y = PMA_PreModelling_Train_KNN$Log_Solubility,
                 method = "knn",
                 tuneGrid = KNN_Grid,
                 trControl = KFold_Control,
                 preProc = c("center", "scale"))

##################################
# Reporting the cross-validation results
# for the train set
##################################
KNN_Tune

KNN_Tune$finalModel

KNN_Tune$results

(KNN_Train_RMSE <- KNN_Tune$results[KNN_Tune$results$k==KNN_Tune$bestTune$k,
                 c("RMSE")])
(KNN_Train_Rsquared <- KNN_Tune$results[KNN_Tune$results$k==KNN_Tune$bestTune$k,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
# model does not support variable importance measurement

##################################
# Independently evaluating the model
# on the test set
##################################
KNN_Test <- data.frame(KNN_Observed = PMA_PreModelling_Test$Log_Solubility,
                         KNN_Predicted = predict(KNN_Tune,
                         PMA_PreModelling_Test_KNN[,!names(PMA_PreModelling_Test_KNN) %in% c("Log_Solubility")]))

KNN_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(KNN_Test_Metrics <- postResample(KNN_Test[,2], KNN_Test[,1]))
(KNN_Test_RMSE <- KNN_Test_Metrics[1])
(KNN_Test_Rsquared <- KNN_Test_Metrics[2])

```

###  1.5.12 Classification and Regression Trees (CART)
|
| **[A]** The classification and regression trees model from the  <mark style="background-color: #CCECFF">**rpart**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">cp</span> = complexity parameter threshold made to vary across a range of values equal to 0.001 to 0.020
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves cp=0.001
|      **[C.2]** Root-Mean-Square Error = 0.94904
|      **[C.3]** R-Squared = 0.78141
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">HydrophilicFactor</span> variable (numeric)
|      **[D.4]** <span style="color: #FF0000">NumBonds</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">SurfaceArea1</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.91946
|      **[E.2]** R-Squared = 0.80707
|
```{r section_1.5.12, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_CART <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_CART$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
CART_Grid = data.frame(cp = c(0.001, 0.005, 0.010, 0.015, 0.020))

##################################
# Running the classification and regression trees model
# by setting the caret method to 'rpart'
##################################
set.seed(12345678)
CART_Tune <- train(x = PMA_PreModelling_Train_CART[,!names(PMA_PreModelling_Train_CART) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_CART$Log_Solubility,
                 method = "rpart",
                 tuneGrid = CART_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
CART_Tune

CART_Tune$finalModel

CART_Tune$results

(CART_Train_RMSE <- CART_Tune$results[CART_Tune$results$cp==CART_Tune$bestTune$cp,
                 c("RMSE")])
(CART_Train_Rsquared <- CART_Tune$results[CART_Tune$results$cp==CART_Tune$bestTune$cp,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
CART_VarImp <- varImp(CART_Tune, scale = TRUE)
plot(CART_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Classification and Regression Trees",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
CART_Test <- data.frame(CART_Observed = PMA_PreModelling_Test$Log_Solubility,
                       CART_Predicted = predict(CART_Tune, 
                       PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

CART_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(CART_Test_Metrics <- postResample(CART_Test[,2], CART_Test[,1]))
(CART_Test_RMSE <- CART_Test_Metrics[1])
(CART_Test_Rsquared <- CART_Test_Metrics[2])

```

###  1.5.13 Conditional Inference Trees (CTREE)
|
| **[A]** The conditional inference trees model from the  <mark style="background-color: #CCECFF">**party**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">mincriterion</span> = 1-p-value threshold made to vary across a range of values equal to 0.75 to 0.99
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves mincriterion=0.75
|      **[C.2]** Root-Mean-Square Error = 0.95704
|      **[C.3]** R-Squared = 0.77940
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">FP076</span> variable (factor)
|      **[D.4]** <span style="color: #FF0000">NumMultBonds</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">NumHalogen</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 1.11601
|      **[E.2]** R-Squared = 0.73109
|
```{r section_1.5.13, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_CTREE <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_CTREE$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
CTREE_Grid = data.frame(mincriterion = sort(c(0.95, seq(0.75, 0.99, length = 2))))

##################################
# Running the conditional inference trees model
# by setting the caret method to 'ctree'
##################################
set.seed(12345678)
CTREE_Tune <- train(x = PMA_PreModelling_Train_CTREE[,!names(PMA_PreModelling_Train_CTREE) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_CTREE$Log_Solubility,
                 method = "ctree",
                 tuneGrid = CTREE_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
CTREE_Tune

CTREE_Tune$finalModel

CTREE_Tune$results

(CTREE_Train_RMSE <- CTREE_Tune$results[CTREE_Tune$results$mincriterion==CTREE_Tune$bestTune$mincriterion,
                 c("RMSE")])
(CTREE_Train_Rsquared <- CTREE_Tune$results[CTREE_Tune$results$mincriterion==CTREE_Tune$bestTune$mincriterion,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
CTREE_VarImp <- varImp(CTREE_Tune, scale = TRUE)
plot(CTREE_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Conditional Inference Trees",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
CTREE_Test <- data.frame(CTREE_Observed = PMA_PreModelling_Test$Log_Solubility,
                       CTREE_Predicted = predict(CTREE_Tune, 
                       PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

CTREE_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(CTREE_Test_Metrics <- postResample(CTREE_Test[,2], CTREE_Test[,1]))
(CTREE_Test_RMSE <- CTREE_Test_Metrics[1])
(CTREE_Test_Rsquared <- CTREE_Test_Metrics[2])

```


###  1.5.14 Random Forest (RF)
|
| **[A]** The random forest model from the  <mark style="background-color: #CCECFF">**randomForest**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package.
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">mtry</span> = number of randomly selected predictors made to vary across a range of values equal to 10 to 100
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves mtry=75
|      **[C.2]** Root-Mean-Square Error = 0.65419
|      **[C.3]** R-Squared = 0.90098
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">HydroPhilicFactor</span> variable (numeric)
|      **[D.4]** <span style="color: #FF0000">SurfaceArea1</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">NumRotBonds</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.65571
|      **[E.2]** R-Squared = 0.90057
|
```{r section_1.5.14, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_RF <- PMA_PreModelling_Train

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_RF$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
RF_Grid = data.frame(mtry = c(25,75,125))

##################################
# Running the random forest model
# by setting the caret method to 'rf'
##################################
set.seed(12345678)
RF_Tune <- train(x = PMA_PreModelling_Train_RF[,!names(PMA_PreModelling_Train_RF) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_RF$Log_Solubility,
                 method = "rf",
                 tuneGrid = RF_Grid,
                 ntree = 100,
                 importance = TRUE,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
RF_Tune

RF_Tune$finalModel

RF_Tune$results

(RF_Train_RMSE <- RF_Tune$results[RF_Tune$results$mtry==RF_Tune$bestTune$mtry,
                 c("RMSE")])
(RF_Train_Rsquared <- RF_Tune$results[RF_Tune$results$mtry==RF_Tune$bestTune$mtry,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
RF_VarImp <- varImp(RF_Tune, scale = TRUE)
plot(RF_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Random Forest",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
RF_Test <- data.frame(RF_Observed = PMA_PreModelling_Test$Log_Solubility,
                       RF_Predicted = predict(RF_Tune, 
                       PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

RF_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(RF_Test_Metrics <- postResample(RF_Test[,2], RF_Test[,1]))
(RF_Test_RMSE <- RF_Test_Metrics[1])
(RF_Test_Rsquared <- RF_Test_Metrics[2])

```

###  1.5.15 Cubist (CUB)
|
| **[A]** The cubist model from the  <mark style="background-color: #CCECFF">**Cubist**</mark> package was implemented through the <mark style="background-color: #CCECFF">**caret**</mark> package. 
|
| **[B]** The model contains 2 hyperparameters:
|      **[B.1]** <span style="color: #FF0000">committees</span> = number of committees made to vary across a range of values equal to 1 to 100
|      **[B.2]** <span style="color: #FF0000">neighbors</span> = number of neighbors made to vary across a range of values equal to 0 to 9
|
| **[C]** The cross-validated model performance of the final model is summarized as follows:
|      **[C.1]** Final model configuration involves committees=100 and neighbors=9
|      **[C.2]** Root-Mean-Square Error = 0.55645
|      **[C.3]** R-Squared = 0.92538
|
| **[D]** The model allows for ranking of predictors in terms of variable importance. The top-performing predictors in the model are as follows:
|      **[D.1]** <span style="color: #FF0000">NumCarbon</span> variable (numeric)
|      **[D.2]** <span style="color: #FF0000">SurfaceArea1</span> variable (numeric)
|      **[D.3]** <span style="color: #FF0000">MolWeight</span> variable (numeric)
|      **[D.4]** <span style="color: #FF0000">SurfaceArea2</span> variable (numeric)
|      **[D.5]** <span style="color: #FF0000">NumMultBonds</span> variable (numeric)
|
| **[E]** The independent test model performance of the final model is summarized as follows:
|      **[E.1]** Root-Mean-Square Error = 0.63809
|      **[E.2]** R-Squared = 0.90717
|
```{r section_1.5.15, warning=FALSE, message=FALSE}
##################################
# Transforming factor predictors
# as required by the nature of the model
##################################
# Creating a local object
# for the train and test sets
##################################
PMA_PreModelling_Train_CUB <- as.data.frame(lapply(PMA_PreModelling_Train, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Train_CUB)

PMA_PreModelling_Test_CUB <- as.data.frame(lapply(PMA_PreModelling_Test, function(x) as.numeric(as.character(x))))
dim(PMA_PreModelling_Test_CUB)

##################################
# Creating consistent fold assignments 
# for the 10-Fold Cross Validation process
##################################
set.seed(12345678)
KFold_Indices <- createFolds(PMA_PreModelling_Train_CUB$Log_Solubility,
                             k = 10,
                             returnTrain=TRUE)
KFold_Control <- trainControl(method="cv",
                              index=KFold_Indices)

##################################
# Setting the conditions
# for hyperparameter tuning
##################################
CUB_Grid = expand.grid(committees = c(1:10, 20, 50, 75, 100), 
                      neighbors = c(0, 1, 5, 9))

##################################
# Running the cubist model
# by setting the caret method to 'cubist'
##################################
set.seed(12345678)
CUB_Tune <- train(x = PMA_PreModelling_Train_CUB[,!names(PMA_PreModelling_Train_CUB) %in% c("Log_Solubility")], 
                 y = PMA_PreModelling_Train_CUB$Log_Solubility,
                 method = "cubist",
                 tuneGrid = CUB_Grid,
                 trControl = KFold_Control)

##################################
# Reporting the cross-validation results
# for the train set
##################################
CUB_Tune

CUB_Tune$finalModel

CUB_Tune$results

(CUB_Train_RMSE <- CUB_Tune$results[CUB_Tune$results$committees==CUB_Tune$bestTune$committees &
                              CUB_Tune$results$neighbors==CUB_Tune$bestTune$neighbors,
                 c("RMSE")])
(CUB_Train_Rsquared <- CUB_Tune$results[CUB_Tune$results$committees==CUB_Tune$bestTune$committees &
                              CUB_Tune$results$neighbors==CUB_Tune$bestTune$neighbors,
                 c("Rsquared")])

##################################
# Identifying and plotting the
# best model predictors
##################################
CUB_VarImp <- varImp(CUB_Tune, scale = TRUE)
plot(CUB_VarImp, 
     top=25, 
     scales=list(y=list(cex = .95)),
     main="Ranked Variable Importance : Cubist",
     xlab="Scaled Variable Importance Metrics",
     ylab="Predictors",
     cex=2,
     origin=0,
     alpha=0.45)

##################################
# Independently evaluating the model
# on the test set
##################################
CUB_Test <- data.frame(CUB_Observed = PMA_PreModelling_Test$Log_Solubility,
                       CUB_Predicted = predict(CUB_Tune, 
                       PMA_PreModelling_Test[,!names(PMA_PreModelling_Test) %in% c("Log_Solubility")]))

CUB_Test

##################################
# Reporting the independent evaluation results
# for the test set
##################################
(CUB_Test_Metrics <- postResample(CUB_Test[,2], CUB_Test[,1]))
(CUB_Test_RMSE <- CUB_Test_Metrics[1])
(CUB_Test_Rsquared <- CUB_Test_Metrics[2])

```

##  1.6 Model Evaluation Summary
|
| Model performance comparison:
|
| **[A]** The models which demonstrated the best and most consistent r-squared and RMSE metrics are as follows:
|      **[A.1]** CUB: Cubist (<mark style="background-color: #CCECFF">**Cubist**</mark> package)
|             **[A.1.1]** Cross-Validation R-Squared = 0.92538, Test R-Squared = 0.90717 
|             **[A.1.2]** Cross-Validation RMSE = 0.55645, Test RMSE = 0.63809
|      **[A.2]** SVM_R: Support Vector Machine - Radial Basis Function Kernel (<mark style="background-color: #CCECFF">**kernlab**</mark> package)
|             **[A.2.1]** Cross-Validation R-Squared = 0.91551, Test R-Squared = 0.90989 
|             **[A.2.2]** Cross-Validation RMSE = 0.59505, Test RMSE = 0.62742
|      **[A.3]** SVM_R: Support Vector Machine - Polynomial Kernel (<mark style="background-color: #CCECFF">**kernlab**</mark> package)
|             **[A.3.1]** Cross-Validation R-Squared = 0.91167, Test R-Squared = 0.90623 
|             **[A.3.2]** Cross-Validation RMSE = 0.59505, Test RMSE = 0.63778
|      **[A.4]** RF: Random Forest (<mark style="background-color: #CCECFF">**randomForest**</mark> package)
|             **[A.4.1]** Cross-Validation R-Squared = 0.90098, Test R-Squared = 0.90057 
|             **[A.4.2]** Cross-Validation RMSE = 0.65419, Test RMSE = 0.65571
|
```{r section_1.6, warning=FALSE, message=FALSE}
##################################
# Consolidating all evaluation results
# for the train and test sets
# using the r-squared metric
##################################
Model <- c('LR','PLR_R','PLR_L','PLR_E','PCR','PLS','AVNN','MARS','SVM_R','SVM_P','KNN','CART','CTREE','RF','CUB',
           'LR','PLR_R','PLR_L','PLR_E','PCR','PLS','AVNN','MARS','SVM_R','SVM_P','KNN','CART','CTREE','RF','CUB')

Set <- c(rep('Cross-Validation',15),rep('Test',15))

R_Squared <- c(LR_Train_Rsquared,PLR_R_Train_Rsquared,PLR_L_Train_Rsquared,PLR_E_Train_Rsquared,PCR_Train_Rsquared,
               PLS_Train_Rsquared,AVNN_Train_Rsquared,MARS_Train_Rsquared,SVM_R_Train_Rsquared,SVM_P_Train_Rsquared,
               KNN_Train_Rsquared,CART_Train_Rsquared,CTREE_Train_Rsquared,RF_Train_Rsquared,CUB_Train_Rsquared,
               LR_Test_Rsquared,PLR_R_Test_Rsquared,PLR_L_Test_Rsquared,PLR_E_Test_Rsquared,PCR_Test_Rsquared,
               PLS_Test_Rsquared,AVNN_Test_Rsquared,MARS_Test_Rsquared,SVM_R_Test_Rsquared,SVM_P_Test_Rsquared,
               KNN_Test_Rsquared,CART_Test_Rsquared,CTREE_Test_Rsquared,RF_Test_Rsquared,CUB_Test_Rsquared)

R_Squared_Summary <- as.data.frame(cbind(Model,Set,R_Squared))

R_Squared_Summary$R_Squared <- as.numeric(as.character(R_Squared_Summary$R_Squared))
R_Squared_Summary$Set <- factor(R_Squared_Summary$Set,
                                        levels = c("Cross-Validation",
                                                   "Test"))
R_Squared_Summary$Model <- factor(R_Squared_Summary$Model,
                                        levels = c("LR",
                                                   "PLR_R",
                                                   "PLR_L",
                                                   "PLR_E",
                                                   "PCR",
                                                   "PLS",
                                                   "AVNN",
                                                   "MARS",
                                                   "SVM_R",
                                                   "SVM_P",
                                                   "KNN",
                                                   "CART",
                                                   "CTREE",
                                                   "RF",
                                                   "CUB"))

print(R_Squared_Summary, row.names=FALSE)

(R_Squared_Plot <- dotplot(Model ~ R_Squared,
                           data = R_Squared_Summary,
                           groups = Set,
                           main = "Regression Model Performance Comparison",
                           ylab = "Model",
                           xlab = "R-Squared",
                           auto.key = list(adj = 1),
                           type = c("p", "h"),       
                           origin = 0,
                           alpha = 0.45,
                           pch = 16,
                           cex = 2))

##################################
# Consolidating all evaluation results
# for the train and test sets
# using the rmse metric
##################################
Model <- c('LR','PLR_R','PLR_L','PLR_E','PCR','PLS','AVNN','MARS','SVM_R','SVM_P','KNN','CART','CTREE','RF','CUB',
           'LR','PLR_R','PLR_L','PLR_E','PCR','PLS','AVNN','MARS','SVM_R','SVM_P','KNN','CART','CTREE','RF','CUB')

Set <- c(rep('Cross-Validation',15),rep('Test',15))

RMSE<- c(LR_Train_RMSE,PLR_R_Train_RMSE,PLR_L_Train_RMSE,PLR_E_Train_RMSE,PCR_Train_RMSE,
               PLS_Train_RMSE,AVNN_Train_RMSE,MARS_Train_RMSE,SVM_R_Train_RMSE,SVM_P_Train_RMSE,
               KNN_Train_RMSE,CART_Train_RMSE,CTREE_Train_RMSE,RF_Train_RMSE,CUB_Train_RMSE,
               LR_Test_RMSE,PLR_R_Test_RMSE,PLR_L_Test_RMSE,PLR_E_Test_RMSE,PCR_Test_RMSE,
               PLS_Test_RMSE,AVNN_Test_RMSE,MARS_Test_RMSE,SVM_R_Test_RMSE,SVM_P_Test_RMSE,
               KNN_Test_RMSE,CART_Test_RMSE,CTREE_Test_RMSE,RF_Test_RMSE,CUB_Test_RMSE)

RMSE_Summary <- as.data.frame(cbind(Model,Set,RMSE))

RMSE_Summary$RMSE<- as.numeric(as.character(RMSE_Summary$RMSE))
RMSE_Summary$Set <- factor(RMSE_Summary$Set,
                           levels = c("Cross-Validation",
                                      "Test"))
RMSE_Summary$Model <- factor(RMSE_Summary$Model,
                             levels = c("LR",
                                        "PLR_R",
                                        "PLR_L",
                                        "PLR_E",
                                        "PCR",
                                        "PLS",
                                        "AVNN",
                                        "MARS",
                                        "SVM_R",
                                        "SVM_P",
                                        "KNN",
                                        "CART",
                                        "CTREE",
                                        "RF",
                                        "CUB"))

print(RMSE_Summary, row.names=FALSE)

(RMSE_Plot <- dotplot(Model ~ RMSE,
                      data = RMSE_Summary,
                      groups = Set,
                      main = "Regression Model Performance Comparison",
                      ylab = "Model",
                      xlab = "Root-Mean-Square Error",
                      auto.key = list(adj = 1),
                      type = c("p", "h"),
                      origin = 0,
                      alpha = 0.45,
                      pch = 16,
                      cex = 2))
```

# **2. References**
|
| **[Book]** [Applied Predictive Modeling](http://appliedpredictivemodeling.com/) by Max Kuhn and Kjell Johnson
| **[Book]** [An Introduction to Statistical Learning](https://www.statlearning.com/) by Gareth James, Daniela Witten, Trevor Hastie and Rob Tibshirani
| **[Book]** [Multivariate Data Visualization with R](http://lmdvr.r-forge.r-project.org/figures/figures.html) by Deepayan Sarkar
| **[Book]** [Machine Learning](https://bookdown.org/ssjackson300/Machine-Learning-Lecture-Notes/) by Samuel Jackson
| **[Book]** [Data Modeling Methods](https://bookdown.org/larget_jacob/data-modeling-methods/) by Jacob Larget
| **[R Package]** [AppliedPredictiveModeling](https://cran.r-project.org/web//packages/AppliedPredictiveModeling/AppliedPredictiveModeling.pdf) by Max Kuhn
| **[R Package]** [caret](https://topepo.github.io/caret/index.html) by Max Kuhn
| **[R Package]** [rpart](https://mran.microsoft.com/web/packages/rpart/rpart.pdf) by Terry Therneau and Beth Atkinson
| **[R Package]** [lattice](https://cran.r-project.org/web/packages/lattice/lattice.pdf) by  Deepayan Sarkar
| **[R Package]** [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html/) by Hadley Wickham
| **[R Package]** [moments](https://cran.r-project.org/web/packages/moments/index.html) by Lukasz Komsta and Frederick
| **[R Package]** [skimr](https://cran.r-project.org/web/packages/skimr/skimr.pdf) by  Elin Waring
| **[R Package]** [RANN](https://cran.r-project.org/web/packages/RANN/RANN.pdf) by  Sunil Arya, David Mount, Samuel Kemp and Gregory Jefferis
| **[R Package]** [corrplot](https://cran.r-project.org/web/packages/corrplot/corrplot.pdf) by Taiyun Wei
| **[R Package]** [tidyverse](https://cran.r-project.org/web/packages/tidyverse/tidyverse.pdf) by Hadley Wickham
| **[R Package]** [lares](https://cran.rstudio.com/web/packages/lares/lares.pdf) by Bernardo Lares
| **[R Package]** [DMwR](https://mran.microsoft.com/snapshot/2016-05-02/web/packages/DMwR/DMwR.pdf) by Luis Torgo
| **[R Package]** [gridExtra](https://cran.r-project.org/web/packages/gridExtra/gridExtra.pdf) by Baptiste Auguie and Anton Antonov
| **[R Package]** [rattle](https://cran.r-project.org/web/packages/rattle/rattle.pdf) by Graham Williams
| **[R Package]** [rpart.plot](https://cran.r-project.org/web/packages/rpart.plot/rpart.plot.pdf) by Stephen Milborrow
| **[R Package]** [RColorBrewer](https://cran.r-project.org/web//packages/RColorBrewer/RColorBrewer.pdf) by Erich Neuwirth
| **[R Package]** [stats](https://search.r-project.org/R/refmans/stats/html/00Index.html) by R Core Team
| **[R Package]** [pls](https://cran.r-project.org/web/packages/pls/pls.pdf) by Kristian Hovde Liland
| **[R Package]** [nnet](https://cran.r-project.org/web/packages/nnet/nnet.pdf) by Brian Ripley
| **[R Package]** [elasticnet](https://cran.r-project.org/web/packages/elasticnet/elasticnet.pdf) by Hui Zou
| **[R Package]** [earth](https://cran.r-project.org/web/packages/earth/earth.pdf) by Stephen Milborrow
| **[R Package]** [party](https://cran.r-project.org/web/packages/party/party.pdf) by Torsten Hothorn
| **[R Package]** [kernlab](https://cran.r-project.org/web/packages/kernlab/kernlab.pdf) by Alexandros Karatzoglou
| **[R Package]** [randomForest](https://cran.r-project.org/web/packages/randomForest/randomForest.pdf) by Andy Liaw
| **[R Package]** [Cubist](https://cran.r-project.org/web/packages/Cubist/Cubist.pdf) by Max Kuhn
| **[Article]** [The caret Package](https://topepo.github.io/caret/index.html) by Max Kuhn
| **[Article]** [A Short Introduction to the caret Package](https://cran.r-project.org/web/packages/caret/vignettes/caret.html) by Max Kuhn
| **[Article]** [Caret Package  A Practical Guide to Machine Learning in R](https://www.machinelearningplus.com/machine-learning/caret-package/#:~:text=Caret%20is%20short%20for%20Classification%20And%20REgression%20Training.,track%20of%20which%20algorithm%20resides%20in%20which%20package.) by Selva Prabhakaran
| **[Article]** [Tuning Machine Learning Models Using the Caret R Package](https://machinelearningmastery.com/tuning-machine-learning-models-using-the-caret-r-package/) by Jason Brownlee
| **[Article]** [Lattice Graphs](http://www.sthda.com/english/wiki/lattice-graphs) by STHDA Team
| **[Article]** [A Tour of Machine Learning Algorithms](https://machinelearningmastery.com/a-tour-of-machine-learning-algorithms/) by Jason Brownlee
| **[Article]** [An Introduction to Multivariate Adaptive Regression Spliness](https://www.statology.org/multivariate-adaptive-regression-splines/) by Statology Team
| **[Article]** [An Introduction to Principal Components Regression](https://www.statology.org/principal-components-regression/) by Statology Team
| **[Article]** [An Introduction to Partial Least Squares](https://www.statology.org/partial-least-squares/) by Statology Team
| **[Article]** [Cubist Regression Models](https://cran.r-project.org/web/packages/Cubist/vignettes/cubist.html) by Max Kuhn
| **[Course]** [Applied Data Mining and Statistical Learning](https://online.stat.psu.edu/stat508/) by Penn State Eberly College of Science
| **[Course]** [Regression Methods](https://online.stat.psu.edu/stat501/) by Penn State Eberly College of Science
| **[Course]** [Applied Regression Analysis](https://online.stat.psu.edu/stat462/) by Penn State Eberly College of Science
| **[Course]** [Applied Data Mining and Statistical Learning](https://online.stat.psu.edu/stat508/) by Penn State Eberly College of Science
|
|
|
|